%\author{FirstName Surname}
\twoauthors{Georg-Christoph Haas}{Stephanie Eckman}
%\threeauthors{FirstName Surname}{FirstNameSurname}{FirstName Surname}
% etc.

\twoaffiliations{Institute for Employment Research}{RTI International \\ Institute for Employment Research}

\leftheader{Haas and Eckman} 
\rightheader{The Title or a short title}

\title{How does Asking for Linkage Consent in the Beginning of the Questionnaire Affect Data Quality?}

\authornote{Contact information:
FirstName Name, Affiliation, Postal addres (email)}

\abstract{Through combining survey data with administrative data researchers additional research is possible. Usually researchers need the consent of each respondent which shall be linked to do so. Naturally we want to obtain the highest consent rates to have the most possible power. It is best practice asking the linkage consent question at the beginning of the questionnaire (Sakshaug et al. 2013). However, if asking this question at the beginning the linkage consent question may have an impact on respondents’ response behavior because they may have the feeling of being observed. We hypothesized that this feeling leads to two possible outcomes. First, respondents effort to answer questions increases because they fear consequences if survey data differs from administrative data (better-respondents hypotheses). Second, respondents may become more satisficing in answering questions because they believe all important information about her/him is covered through the administrative data and survey responses can be corrected afterwards.
For our analyses we use a German Survey that asked one half of respondents at the beginning and the other half at the end. Since both groups suffer from selection bias, we use the entropy balance technique to simulate a randomized distribution over both groups and calculating average treatment effects on measurement error and item nonresponse. Also we compare the effects for a CATI and web mode. Our results show that no effects in either of these two modes occur, when asking the administrative data linkage question at the beginning of the interview.
}

\keywords{measurement error; linkage consent; data quality}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "srm_main"
%%% End: 

